---
title: "PMI Questions"
author: "Chathura_Perera"
date: "2022-11-14"
output: html_document
---

```{r}
library(xlsx)
library(openxlsx)
library(writexl)
library(dplyr)
library(tidyr)
library(stringr)
library(lubridate)
library(formattable)
library(readxl)
library(imputeTS)
library(ggplot2)
library(caret)

```




```{r}

pathlocal <- r"(C:\Users\cperera\OneDrive - Digicorner\Desktop\rgu\item.xlsx)" #pathlocal
pathonline <- r"(C:\Users\cperera\OneDrive - Digicorner\Desktop\rgu\item.xlsx)" #pathonline
otherpath <- r"(C:\Users\94761\Downloads\item.xlsx)"

hr <- readWorkbook(otherpath,sheet = "Data",startRow = 1,colNames = TRUE,rowNames = FALSE,detectDates = TRUE)


```


```{r}


hr$Leaver <- as.factor(hr$Leaver)


```



```{r}
head(hr)
```



```{r}

names(hr)

```
# Does gender impact attrition
```{r}


hr |> group_by(Function,Leaver) |> summarise(frequencies = n()) |> 
mutate(relative_proportion = 100*round(frequencies / sum(frequencies), 3),Cumulative_sum=cumsum(frequencies),Cumulative_prop = 100*round(Cumulative_sum / sum(frequencies),3) ) |> filter(Leaver==1) |> 
  #rounded to 3 decimal points 
arrange(desc(relative_proportion))

```
# marketing , external affrairs 
# comments - 
# Insight  - female attition is higher
# Add a visual 



```{r}

hr |> group_by(Employee.Group,Leaver) |> summarise(frequencies = n()) |> 
mutate(relative_proportion = 100*round(frequencies / sum(frequencies), 3),Cumulative_sum=cumsum(frequencies),Cumulative_prop = 100*round(Cumulative_sum / sum(frequencies),3) ) |> 
  #rounded to 3 decimal points 
arrange(Cumulative_sum)


```
# Insight to prove temps are leaving more frequently
# HBR maybe ?? 





```{r}

hr |> group_by(Country,Leaver) |> summarise(frequencies = n()) |> 
mutate(relative_proportion = 100*round(frequencies / sum(frequencies), 3),Cumulative_sum=cumsum(frequencies),Cumulative_prop = 100*round(Cumulative_sum / sum(frequencies),3) ) |> filter(Leaver==1) |> 
  #rounded to 3 decimal points 
arrange(desc(relative_proportion))

```

# line manger rating score
```{r}

hr |> group_by(Line.manager.rating.survey.score,Leaver) |> summarise(frequencies = n()) |> 
mutate(relative_proportion = 100*round(frequencies / sum(frequencies), 3),Cumulative_sum=cumsum(frequencies),Cumulative_prop = 100*round(Cumulative_sum / sum(frequencies),3) ) |> filter(Leaver==1) |> 
  #rounded to 3 decimal points 
arrange(desc(relative_proportion))

```


# nothing significant 
```{r}
ggplot(hr, aes(x=Gender,y=Hourly.Rate))+
geom_boxplot(alpha = 0.1,aes(color=Gender))+
geom_jitter(size = 1.5,alpha=0.12,aes(color=Gender))+
ggtitle("Hourly rate  distribution by Gender")+
theme(plot.title = element_text(face="bold",hjust = 0.5))
```






```{r}

#Recognition.Survey.score
#Line.manager.rating.survey.score
hr |> filter(Leaver==1) |> 
ggplot( aes(x=Recognition.Survey.score,y=Line.manager.rating.survey.score))+
geom_jitter(aes(color=Leaver))+
ggtitle("Salary hike vs Performance Rating")+
labs(x="Performance Rating", y="Salary Hike")+ 
theme(plot.title = element_text(face="bold",hjust = 0.5))


```

```{r}

hr |> filter(Leaver==1) |> 

ggplot( aes(x=Career.opportuntities.Survey.score,y=Time.in.latest.role))+
geom_jitter(aes(color=Leaver),alpha=0.6)

```


# there is connection between not having direct repots and attrtion 
```{r}


hr |> 

ggplot( aes(x=eNPS.Score,y=Actual.monthly.pay))+
geom_jitter(aes(color=Leaver),alpha=0.6)



```

```{r}

ggplot(hr, aes(x=Gender,y=Hourly.Rate))+
geom_boxplot(alpha = 0.1,aes(color=Gender))+
geom_jitter(size = 1.5,alpha=0.12,aes(color=Gender))+
ggtitle("Hourly rate  distribution by Gender")+
theme(plot.title = element_text(face="bold",hjust = 0.5))


```

## Summary of model Application ## 
## Insight = how did that transform into the model application 
## this processing will make sure that inisghts were detected and used 


# Attrition prediction 

```{r}
hr <- hr |> select(-Employee.ID)

set.seed(123)


index <- sample(2, nrow(hr), replace = TRUE, prob = c(0.7, 0.3))
train <- hr[index==1,]
test <- hr[index==2,]


```


```{r}

train %>% 
group_by(Leaver) %>% 
summarise(count =n()) %>% 
mutate(relative_proportion = 100*round(count/sum(count), 3))

```
# Observe that non attritions are a little more that 5 times of attritions



```{r}


trainup <- 
train %>% 
upSample( train$Leaver) %>% 
select(-c("Leaver")) %>% 
rename(Leaver=Class)

```



```{r}

trainup %>% 
group_by(Leaver) %>% 
summarise(count =n()) %>% 
mutate(relative_proportion = 100*round(count/sum(count), 3))

```

```{r}
set.seed(123)
xGB.modelbasic <- 
train(Leaver ~ ., 
data=trainup,   
metric = "Accuracy",
method="xgbTree",
verbosity = 0,
preProcess=c("center","scale"))
```



```{r}
prediction = predict(xGB.modelbasic, newdata=test)
confusionMatrix(prediction,test$Leaver)
```

```{r}
varImp(xGB.modelbasic)
```

```{r}
plot(xGB.modelbasic)
```


```{r}

control = trainControl(method = "repeatedcv", number=10,repeats = 5)

```

```{r}

xgbGrid01 <- expand.grid(nrounds = c(20,40,60,80),
                       max_depth = c(10,15,20,25),
                       eta = c(0.1,0.2,0.4),
                       gamma = 0,
                       colsample_bytree = 0.7,
                       min_child_weight = 1,
                       subsample = c(0.6,0.8,1))


```

```{r}
hr$model_weights <- ifelse(hr$Leaver == "Yes",
                        table(hr$Leaver)[1]/nrow(hr),
                        table(hr$Leaver)[2]/nrow(hr))


set.seed(123)
index <- sample(2, nrow(hr), replace = TRUE, prob = c(0.7, 0.3))
train <- hr[index==1,]
test <- hr[index==2,]


trainup <- 
train %>% 
upSample( train$Leaver) %>% 
select(-c("Leaver")) %>% 
rename(Leaver=Class)



```

```{r}
set.seed(123)
xGB.modelenhanced <- 
train(Leaver ~ ., 
data=trainup,   
metric = "Accuracy",
method="xgbTree",
verbosity = 0,
weights = model_weights , #adding model weights 
preProcess=c("center","scale"),
trControl = control, #cross validation 
tuneGrid = xgbGrid01) # hyperparameter tuning 
```
```{r}
prediction = predict(xGB.modelenhanced, newdata=test)
confusionMatrix(prediction,test$Leaver) 
```



```{r}
varImp(xGB.modelenhanced)
```


```{r}

```



